---
title: "Testing OptiFit"
author: "Courtney R Armour"
date: "1/31/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)

library(tidyverse)

```
### MCC OptiFit vs OptiClust 

We've been running OptiFit using 80% of samples as the reference. I wanted to see if there would be any change if a higher % of samples was used in the reference. Here I am comparing the MCC scores from 80% (8020) of samples as reference to 90% (9010) of samples as reference. Additionally, when chatting with Kelly about OptiFit she mentioned that she did not do any subsampling in her OptiFit analysis. I ran an addition set where I removed all the subsampling steps and the MCC is plotted below. 

```{r mcc}
presubsamp_merged_mcc <- read_csv("data/presubsamp_merged_MCC.csv") %>% 
  mutate(label = case_when(algorithm == "opticlust" ~ algorithm,
                           TRUE ~ paste0(algorithm,"_",subset))) %>% 
  mutate(Group=case_when(algorithm == "opticlust" ~ "OptiClust",
                         subset == "train" ~ "OptiFit Reference",
                         state == "fit" ~ "Optifit Fit w/o Reference",
                         TRUE ~ "Optifit Fit w/ Reference")) %>% 
  mutate(label = factor(label,levels=c("opticlust","optifit_train","optifit_test"),
                        labels = c("OptiClust","OptiFit Train","OptiFit Test"))) %>% 
  mutate(data_split = "8020 - presubsamp")
```

```{r mcc}
postsubsamp_merged_mcc <- read_csv("data/postsubsamp_merged_MCC.csv") %>% 
  mutate(label = case_when(algorithm == "opticlust" ~ algorithm,
                           TRUE ~ paste0(algorithm,"_",subset))) %>% 
  mutate(Group=case_when(algorithm == "opticlust" ~ "OptiClust",
                         subset == "train" ~ "OptiFit Reference",
                         state == "fit" ~ "Optifit Fit w/o Reference",
                         TRUE ~ "Optifit Fit w/ Reference")) %>% 
  mutate(label = factor(label,levels=c("opticlust","optifit_train","optifit_test"),
                        labels = c("OptiClust","OptiFit Train","OptiFit Test"))) %>% 
  mutate(data_split = "8020 - postsubsamp")
```

```{r mcc_9010}
merged_mcc_9010 <- read_csv("data/9010split_merged_MCC.csv") %>% 
  mutate(label = case_when(algorithm == "opticlust" ~ algorithm,
                           TRUE ~ paste0(algorithm,"_",subset))) %>% 
  mutate(Group=case_when(algorithm == "opticlust" ~ "OptiClust",
                         subset == "train" ~ "OptiFit Reference",
                         state == "fit" ~ "Optifit Fit w/o Reference",
                         TRUE ~ "Optifit Fit w/ Reference")) %>% 
  mutate(label = factor(label,levels=c("opticlust","optifit_train","optifit_test"),
                        labels = c("OptiClust","OptiFit Train","OptiFit Test"))) %>% 
  mutate(data_split = "9010 - presubsamp")
```


```{r nosubsamp}
merged_mcc_nosubsamp <- read_csv("data/nosubsamp_merged_MCC.csv") %>% 
  mutate(label = case_when(algorithm == "opticlust" ~ algorithm,
                           TRUE ~ paste0(algorithm,"_",subset))) %>% 
  mutate(Group=case_when(algorithm == "opticlust" ~ "OptiClust",
                         subset == "train" ~ "OptiFit Reference",
                         state == "fit" ~ "Optifit Fit w/o Reference",
                         TRUE ~ "Optifit Fit w/ Reference")) %>% 
  mutate(label = factor(label,levels=c("opticlust","optifit_train","optifit_test"),
                        labels = c("OptiClust","OptiFit Train","OptiFit Test"))) %>% 
  mutate(data_split = "No Subsample")

```


```{r plot}

bind_rows(presubsamp_merged_mcc,merged_mcc_9010) %>% 
  bind_rows(merged_mcc_nosubsamp) %>% 
  bind_rows(postsubsamp_merged_mcc) %>% 
  ggplot(aes(x=label,y=mcc,color=Group)) +
    geom_jitter(height=0,width=0.2,alpha=0.5,size=3) +
    theme_bw() +
    xlab("") + ylab("MCC") +
    theme(panel.grid.major.y = element_line(color="grey85"),
          panel.grid.minor.y = element_line(color="grey85"),
          panel.grid.major.x = element_blank(),
          legend.position = "top",
          axis.text.x = element_text(angle=45,hjust=1)) +
    #geom_text(aes(label,mcc,label=split),data=subset(merged_mcc, mcc < 0.7),color="black",hjust=0) +
    facet_grid(.~data_split)

```